Road Segment Prioritization for Bicycle Infrastructure

### 2020-11-14 README.Rmd
added latex tables

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. Therefore, the aim of this study is to develop a methodology that determines where cycling infrastructure should be added by accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand in a city will come from, how such demand should be routed on the street network, and how to  consider fair distribution of cycling infrastructure that is in line with egalitarian principles. The city of Manchester is used as a case study, but the methodology is generic enough to be applicable to any city. For routing, it is found that deviating from shortest paths to avoid high-stress road segments can result in acceptable travel time increases for cyclists. Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure. The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one. \\par\\textbf{Keywords:} cycling network, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable"]
---

@@

- Lit Review: I can also organize this thematically and trim down **Planning Cycling Networks** substantially. I would then add some literature in:
  - Calculating Potential Demand
  - Routing
  - Road Segment Prioritization
- Community Detection: Where in the document should this section be?
- Should we use other cities for the sake of comparison
- How to add tables
- Footnotes are sometimes split between pages. How do I keep them on the same page?

@@

The 2015 Paris agreement [@agreement2015paris] was an acknowledgment that fundamental changes to our lives and economies are necessary to reduce the impacts of climate change. The transport sector is one of many sectors facing transformation, with innovation and policy measures leading to the introduction of ‘greener’ private vehicles. But as demand for transport rises, resulting in increased congestion, public and active modes of transport are gaining more recognition. The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; Physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic]. Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country. While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:

@@

The funding on its own is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling. A considerable amount of research has been done towards that end. Segregated cycling infrastructure^[Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators]. Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal]. However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far]. Another defining feature for cycling infrastructure is how well connected it is. Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining]. While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].

As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the whole cycling network. They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions. *Optimization* techniques have been used to propose improvements to cycling networks. @mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget. The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network. The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times. To improve cohesion of the suggested network, a constraint is added so that each link with a bike lane should be connected to at least one destination. Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network. The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network. The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments. The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization]. In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.

These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle]. To solve for continuity, the graph-theoretic concept of *connected components*, has been used. @natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links^[*link* refers to a road segment throughout this research] and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added. They observe that small investments at strategic points have a large impact on connectivity in most cases. The concept of connected components is also at the core of the methodology proposed by @olmos2020data. After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow^[*flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network. The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component. The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.

The problem formulations outlined above look at the network as a whole when attempting to improve it. An alternative approach is to identify the different sub-networks that exist within the larger network, and work on improving each separately. Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them. @akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran. An optimization problem is then formulated to connect nodes within each community with cycling infrastructure. The emphasis is on connectivity within the communities, not between them. @bao2017planning adopt a similar methodology, but use hierarchical clustering to specify the desired number of clusters. They use a greedy network expansion algorithm, where the link with the highest benefit/cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met. The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.

@@

Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network. To avoid this issue, @duthie2014optimization choose to ignore demand completely, and focus on creating a network that connects the entire study area. @olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data. They then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.

@@

```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
```

@@

The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.

**ADD TABLE**

@@

1.  **Cyclist Preference**: Work done by @dill2013four on examining cyclist typologies determined that around 60% of Portland residents fit under the *interested but concerned* category. These were people that enjoyed cycling but avoided it due safety concerns. The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
2.  **Low-Traffic Neighborhoods**: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020]. This includes closing off residential streets to motorized traffic
3.  **Existing Cycling Infrastructure**: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.

The weighting profiles are therefore adjusted to favor less-stressful streets (based on information from Table **REFERENCE THE TABLE**), and roads with existing cycling infrastructure. This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.

**ADD TABLE**

@@

The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different. From Figure \@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester. Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (**Reference the table**), and are therefore part of the Primary Route Network (PRN) [@department2012guidance]. The PRN has the widest, most direct roads on the network, and carries most of the through traffic. This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes [@department2012guidance].

@@

Routing potential cycling demand on a weighted network is more in line with government policy to create LTNs. Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely. This is intentional, as the impedance on these roads is only slightly higher than remaining road types (See Table **Reference the table**). Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.

Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity). When routing using the weighting profile in Table **Reference Table**, we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%. When routing on primary and trunk roads is banned, the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.

@@

Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(what-affects-the-decision-to-cycle), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure. In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].

@@

It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study. It creates a hierarchy of road preference that is grounded in cyclist preference and government plans to create LTNs. Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks , <!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city. More accurate routing could be carried out given the availability of road-level data. In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.

@@

```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="45%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
```

@@

After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments. This cumulative demand (referred to as *flow* is then used as a basis for determining where best to invest in segregated cycling infrastructure. In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(what-affects-the-decision-to-cycle), namely direct and well connected routes.

For this purpose, two algorithms are proposed. Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism. In both algorithms, links^[*link* refers to a road segment] are selected iteratively and the iteration at which each link is added to the solution is recorded. Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.

@@

This algorithm ensures that the resulting network is connected. It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(what-affects-the-decision-to-cycle).

@@

The first algorithm focuses on connectivity and directness, but not on fairly distributing investment. The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(underlying-ethical-principles). This algorithm incorporates the ideal of fair distribution by using community detection to partition the road network.

@@

It is also important to understand how the different highway types contribute to the proposed network. Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table **reference the table**.

@@

<!-- # Conclusions -->

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. Therefore, the aim of this study is to develop a methodology that determines where cycling infrastructure should be added by accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand in a city will come from, how such demand should be routed on the street network, and how to  consider fair distribution of cycling infrastructure that is in line with egalitarian principles. The city of Manchester is used as a case study, but the methodology is generic enough to be applicable to any city. For routing, it is found that deviating from shortest paths to avoid high-stress road segments can result in acceptable travel time increases for cyclists. Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure. The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one. \\par\\textbf{Keywords:} cycling network, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

- Intro: trim down substantially then add some literature in: Calculating Potential Demand / Routing / Road Segment Prioritization
- Community Detection: Where in the document should this section be?
- Should we use other cities for the sake of comparison
- How to add tables properly
- Footnotes are sometimes split between pages. How do I keep them on the same page?

@@

The 2015 Paris agreement [@agreement2015paris] was an acknowledgment that fundamental changes to our lives and economies are necessary to reduce the impacts of climate change. The transport sector is one of many sectors facing transformation, with innovation and policy measures leading to the introduction of ‘greener’ private vehicles. But as demand for transport rises, resulting in increased congestion, public and active modes of transport are also gaining more recognition. The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic]. Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country. While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:

@@

The funding on its own is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.

A considerable amount of research has been done towards that end. Segregated cycling infrastructure^[Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators]. Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal]. However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far]. Another defining feature for cycling infrastructure is how well connected it is. Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining]. While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].

As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the cycling network as a whole. They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions. *Optimization* techniques have been used to propose improvements to cycling networks. @mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget. The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network. The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times. To improve cohesion of the suggested network, a constraint is added so that each link^[*link* refers to a road segment throughout this research] with a bike lane should be connected to at least one destination. Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network. The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network. The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments. The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization]. In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.

These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle]. To solve for continuity, the graph-theoretic concept of *connected components*, has been used. @natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added. They observe that small investments at strategic points have a large impact on connectivity in most cases. The concept of connected components is also at the core of the methodology proposed by @olmos2020data. After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow^[*flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network. The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component. The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.

The problem formulations outlined above look at the network as a whole when attempting to improve it. An alternative approach is to identify the different sub-networks that exist within the larger network, and work on improving each separately. Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them. @akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran. An optimization problem is then formulated to connect nodes within each community with cycling infrastructure. The emphasis is on connectivity within the communities, not between them. @bao2017planning adopt a similar methodology, but use hierarchical clustering to specify the desired number of clusters. They use a greedy network expansion algorithm, where the link with the highest benefit-cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met. The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.

@@

Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network. To avoid this issue, @duthie2014optimization choose to ignore demand completely, and focus on creating a network that connects the entire study area. @olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.

@@

```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="75%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
```

@@

The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. **TALK ABOUT WEIGHTED ROUTING IN OTHER STUDIES **
<!-- This expands on the work of @mauttone2017bicycle, by going beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference. -->

\begin{longtable}{lll}
\caption{OSM Road Types}
\label{table:osmroadtypes}\\
\textbf{OSM Road Type } & \textbf{Description}                                                                                                                                            & \textbf{UK Equivalent}                                                                                                   \endfirsthead 
\hline
Motorway                & \begin{tabular}[c]{@{}l@{}}Road open to high-speed \\vehicles only\end{tabular}                                                                                 & Motorway                                                                                                                 \\* 
\hline
Trunk                   & \multirow{5}{*}{\begin{tabular}[c]{@{}l@{}}Roads that don't \\meet motorway criteria, \\in descending order \\of importance and \\through-traffic\end{tabular}} & A-Roads with Primary Status                                                                                              \\* 
\cline{3-3}
Primary                 &                                                                                                                                                                 & A-Roads with Non-Primary Status                                                                                          \\* 
\cline{3-3}
Secondary               &                                                                                                                                                                 & B-roads                                                                                                                  \\* 
\cline{3-3}
Tertiary                &                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}Classified unnumbered roads \\OR unclassified busy through roads\end{tabular}                 \\* 
\cline{3-3}
Unclassified            &                                                                                                                                                                 & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Unclassified (Intended for \\local traffic - 60\% of UK roads)\end{tabular}}  \\* 
\cline{1-2}
Residential             & Function is purely residential                                                                                                                                  &                                                                                                                          \\ 
\hline
Service                 & \begin{tabular}[c]{@{}l@{}}Road that provides \\access to a facility\end{tabular}                                                                               &                                                                                                                          \\ 
\hline
Track                   & \begin{tabular}[c]{@{}l@{}}Unpaved - suitable for \\two track vehicles (mostly rural)\end{tabular}                                                              &                                                                                                                          \\ 
\hline
Cycleway                & \begin{tabular}[c]{@{}l@{}}Designated cycleway - \\open to cyclists only\end{tabular}                                                                           &                                                                                                                          \\ 
\hline
Path                    & \begin{tabular}[c]{@{}l@{}}Unpaved - open to \\non-motorized traffic only\end{tabular}                                                                          &                                                                                                                          \\
\hline
\end{longtable}

@@

1.  **Cyclist Preference**: Work done by @dill2013four on examining cyclist typologies determined that around 60% of Portland residents fit under the *interested but concerned* category. These were people that enjoyed cycling but avoided it due safety concerns. The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
2.  **Low-Traffic Neighborhoods**: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020]. This includes closing off residential streets to motorized traffic.
3.  **Existing Cycling Infrastructure**: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.

The weighting profiles are therefore adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure. This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.

<!-- **ADD TABLE - THIS IS BASIC** -->
<!-- ```{r, echo=FALSE, message = FALSE} -->
<!-- weight_profiles <- readxl::read_excel("Paper/paper_tables.xlsx", sheet = "Weighting Profile") -->

<!-- knitr::kable(weight_profiles, -->
<!--              caption = "Weighting Profiles") -->
<!-- ``` -->
\begin{longtable}{lrrr}
\caption{Weighting Profiles}
\label{table:weightprofiles}\\
\multirow{2}{*}{\textbf{OSM Road Type}} & \multicolumn{3}{c}{\textbf{Weighting Profile}}                                                 \\*
                                        & \multicolumn{1}{l}{\textit{Unweighted}} & \multicolumn{1}{l}{\textit{Weighted}} & \multicolumn{1}{l}{\textit{Weighted\_2}}  \endfirsthead
Cycleway                                & 1          & 1                                     & 1                                         \\
Path                                    & 1          & 0.9                                   & 0.9                                       \\
Residential                             & 1          & 0.9                                   & 0.9                                       \\
Service                                 & 1          & 0.9                                   & 0.9                                       \\
Tertiary                                & 1          & 0.9                                   & 0.9                                       \\
Track                                   & 1          & 0.9                                   & 0.9                                       \\
Unclassified                            & 1          & 0.9                                   & 0.9                                       \\
Secondary                               & 1          & 0.8                                   & 0.8                                       \\
Primary                                 & 1          & 0.7                                   & \textcolor{red}{0}                        \\
Trunk                                   & 1          & 0.6                                   & \textcolor{red}{0}                        \\
Motorway                                & 1          & \textcolor{red}{0}                    & \textcolor{red}{0}                       
\end{longtable}

@@

The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different. From Figure \@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester. Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (Table \@ref(table:osmroadtypes)), and are therefore part of the Primary Route Network (PRN) [@department2012guidance]. The PRN has the widest, most direct roads on the network, and carries most of the through traffic. This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes [@department2012guidance].

@@

Routing potential cycling demand on a weighted network is more in line with government policy to create LTNs. Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely. This is intentional, as the impedance on these roads is only slightly higher than remaining road types (See Table \@ref(table:weightprofiles)). Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.

Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity). When routing using the weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%. When routing on primary and trunk roads is banned (weighted_2 profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.

@@

Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure. In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].

@@

It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study. It creates a hierarchy of road preference that is grounded in cyclist preference and government plans to create LTNs. Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city. More accurate routing could be carried out given the availability of road-level data. In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.

@@

```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="30%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
```

@@

After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments. This cumulative demand (referred to as *flow* is then used as a basis for determining where best to invest in segregated cycling infrastructure. In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.

For this purpose, two algorithms are proposed. Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism. In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded. Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.

@@

This algorithm ensures that the resulting network is connected. It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(introduction).

@@

The first algorithm focuses on connectivity and directness, but not on fairly distributing investment. The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(introduction). This algorithm incorporates the ideal of fair distribution by using community detection to partition the road network.

@@

It is also important to understand how the different highway types contribute to the proposed network. Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table \@ref(table:weightprofiles).

@@

# Conclusions

The research set out to provide a methodology for designing a cycling network that respects both the needs of the users and the ambitions of the government. The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the census). It should be noted that the recommendations are only as good as the granularity of the available data.  
A core part of the methodology is determining where there is high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand. A routing engine was used to study the effects of limiting access on different road types, and it was found that reducing cyclist flow on roads with high through-traffic resulted in acceptable increases in commuting distances. As a result, a hierarchy of road preference was used to route potential cycling demand. Algorithms were developed to determine where investments in cycling infrastructure should be prioritized. These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.

The research hopes to aid in the current shift towards active transport that is being promoted by the UK government. The success of such a shift is contingent on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.
